---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r knitr, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  echo     = FALSE
)
```

# NBA positions dataset

<!-- badges: start -->
<!-- badges: end -->

This repository contains the NBA positions dataset. The main dataset is designed
to replace the iris dataset and contains basic statistics about 150 NBA Centers,
Point Guards and Shooting Guards from 2017. There is a also a expanded dataset
(`nba_positions_full`) the includes the same statistics for all NBA players in
2017.

## Source

The source file for creating this dataset was obtained from Kaggle at
https://www.kaggle.com/drgilermo/nba-players-stats/data?select=Seasons_Stats.csv
which was originally scraped from [Basketball Reference][basketball-ref].

If you use this dataset please acknowledge the Kaggle dataset and Basketball
Reference (following the [instructions here][basketball-ref-citations]).

## Processing

Here is a brief description of how the dataset was created but see
`nba_positions.R` for details:

1. Read the Kaggle dataset
2. Select rows where `Year == 2017`
3. Select columns containing player, team, games played, positions, turnover
   percentage, rebound percentage, assist percentage and field goal percentage
4. Scale field goal percentage to the range 0-100 to match other statistics
   (this is the `nba_positions_full` dataset)
5. Summarise statistics for players who played on multiple teams in 2017
6. Select Centers, Point Guards and Shooting Guards who played more than 10
   games
7. Cluster the players using k-means with three clusters
8. Select 50 Centers that are most like their cluster, 50 random Point Guards
   and 50 random Shooting Guards

# Summary

Their are two TSV files. The `nba_positions.tsv` file contains the dataset most
similar to the iris data.

* **nba_positions.tsv** 150 selected NBA players with these columns:
  *  `Position` - playing position ("Center", "PointGuard" or "ShootingGuard")
  * `TurnoverPct` - turnover percentage
  * `ReboundPct` - rebound percentage
  * `AssistPct` - assist percentage
  * `FieldGoalPct` - field goal percentage
* **nba_positions_full.tsv** all NBA players in 2017 with these _additional_
  columns:
  * `Player` - player name
  * `Team` - team name
  * `Games` - games played

## Distributions

Example plots showing distributions of the different statistics.

```{r setup}
library(ggplot2)
library(readr)

nba_positions <- read_tsv(
    "nba_positions.tsv",
    col_types = cols(
        Position     = col_character(),
        TurnoverPct  = col_double(),
        ReboundPct   = col_double(),
        AssistPct    = col_double(),
        FieldGoalPct = col_double()
    )
)

theme_set(
    theme_minimal() +
    theme(
        plot.title.position = "plot",
    )
)
```

### Turnover percentage

```{r turnover}
ggplot(nba_positions, aes(x = TurnoverPct, fill = Position)) +
    geom_histogram(bins = 30, position = "identity", alpha = 0.5) +
    scale_fill_manual(values = c("#CB48B7", "#FC440F", "#337357")) +
    scale_x_continuous(
        breaks = seq(5, 95, 5),
        labels = scales::percent_format(scale = 1, accuracy = 1)
    ) +
    labs(
        x     = "Turnover percentage",
        y     = "Frequency",
        title = "Player turnover percentage by position"
    )
```

### Rebound percentage

```{r rebound}
ggplot(nba_positions, aes(x = ReboundPct, fill = Position)) +
    geom_histogram(bins = 30, position = "identity", alpha = 0.5) +
    scale_fill_manual(values = c("#CB48B7", "#FC440F", "#337357")) +
    scale_x_continuous(
        breaks = seq(5, 95, 5),
        labels = scales::percent_format(scale = 1, accuracy = 1)
    ) +
    labs(
        x     = "Rebound percentage",
        y     = "Frequency",
        title = "Player rebound percentage by position"
    )
```

### Assist percentage

```{r assist}
ggplot(nba_positions, aes(x = AssistPct, fill = Position)) +
    geom_histogram(bins = 30, position = "identity", alpha = 0.5) +
    scale_fill_manual(values = c("#CB48B7", "#FC440F", "#337357")) +
    scale_x_continuous(
        breaks = seq(5, 95, 5),
        labels = scales::percent_format(scale = 1, accuracy = 1)
    ) +
    labs(
        x     = "Assist percentage",
        y     = "Frequency",
        title = "Player assist percentage by position"
    )
```

### Field goal percentage

```{r field-goal}
ggplot(nba_positions, aes(x = FieldGoalPct, fill = Position)) +
    geom_histogram(bins = 30, position = "identity", alpha = 0.5) +
    scale_fill_manual(values = c("#CB48B7", "#FC440F", "#337357")) +
    scale_x_continuous(
        breaks = seq(5, 95, 5),
        labels = scales::percent_format(scale = 1, accuracy = 1)
    ) +
    labs(
        x     = "Field goal percentage",
        y     = "Frequency",
        title = "Player field goal percentage by position"
    )
```

[basketball-ref]: http://www.basketball-reference.com/ "Basketball Reference"
[basketball-ref-citations]: https://www.basketball-reference.com/about/contact.html "Basketball Reference Contact"
